A fully labelled proof system for intuitionistic modal logics

### 2020-09-30 FullyLabelled/JLC/introduction.tex
bugfixes in 1 and 2

Standard labelled sequent calculi for logics with
standard Kripke semantics attach to every formla $\fm A$ a label
$\lb x$, witten as $\labels xA$, and additionally use
\emph{relational atoms} of the form $\accs xy$ where 
the binary relation symbol $\rel$
%These calculi work best for logics with
%standard Kripke semantics, as in this case the relation $\rel$ 
is used to encode the accessibility relation in the Kripke models, such that
frame conditions corresponding to the desired logic can be directly
encoded as inference rules.
%\todo{Reviewer 1 complained about the ``best''}
Prominent examples are classical modal
logics and intuitionistic propositional logic, e.g., the frame
condition of transitivity ($\forall \lb {xyz}.\:\accs xy\AND \accs yz\IMP\accs xz$) can be translated into the
inference rule
\begin{equation}
\label{eq:Rtrans}
\vlinf{\rn{trans}}{}{\B, \accs xy, \accs yz, \Left \SEQ \Right}{
	\B, \accs xy, \accs yz, \accs xz, \Left \SEQ \Right}
\end{equation}
where $\B$ stands for a set of relational atoms, and $\Gamma$ and
$\Delta$ for multi-sets of labelled formulas~\cite{negri:vonplato:01}.

@@

	\qquad\qquad

@@

	sequent, we only have that a formula is provable in an extension of $\IK$ beyond paths axioms 

@@

labelled sequent by allowing both, relational atoms of the

@@

of logics that had been treated proof-theoretically before.  

@@

Besides that, there is another pleasant observation to make about our
system. Ordinary labelled sequent systems for intuitionistic modal
logic are single-conclusion~\cite{simpson:phd}. 
%
The same is true for the corresponding nested sequent
systems~\cite{str:fossacs13,marin:str:aiml}. It is possible to express
Maehara style multiple-conclusion systems in nested
sequents~\cite{str:2017maehara}, and therefore also in ordinary
labelled sequents. However, also in these systems there are rules
($\rlabrn\IMP$ and $\rlabrn\BOX$) that force a single-conclusion
premise, even though this is not the case in labelled
systems~\cite{negri:jpl2005} or nested sequents~\cite{fitting:83} for
intuitionistic logic.

>>>>

Standard labelled sequent calculi for logics with
standard Kripke semantics attach to every formla $\fm A$ a label
$\lb x$, witten as $\labels xA$, and additionally use
\emph{relational atoms} of the form $\accs xy$ where 
the binary relation symbol $\rel$
%These calculi work best for logics with
%standard Kripke semantics, as in this case the relation $\rel$ 
is used to encode the accessibility relation in the Kripke models, such that
frame conditions corresponding to the desired logic can be directly
encoded as inference rules.
%\todo{Reviewer 1 complained about the ``best''}
Prominent examples are classical modal
logics and intuitionistic propositional logic, e.g., the frame
condition of transitivity ($\forall \lb {xyz}.\:\accs xy\AND \accs yz\IMP\accs xz$) can be translated into the
inference rule
\begin{equation}
\label{eq:Rtrans}
\vlinf{\rn{trans}}{}{\B, \accs xy, \accs yz, \Left \SEQ \Right}{
	\B, \accs xy, \accs yz, \accs xz, \Left \SEQ \Right}
\end{equation}
where $\B$ stands for a set of relational atoms, and $\Gamma$ and
$\Delta$ for multi-sets of labelled formulas~\cite{negri:vonplato:01}.\lutz{right ref?}

@@

	\hskip10em

@@

	sequent, we only have the result that a formula is provable in an extension of $\IK$ beyond paths axioms 

@@

labelled sequents by allowing both, relational atoms of the

@@

of logics that had not been treated proof-theoretically before.  

@@

Besides that, there is another pleasant observation to make about our
system. It is naturally a multi-conclusion system. Ordinary labelled sequent systems for intuitionistic modal
logic are single-conclusion~\cite{simpson:phd}. 
%
The same is true for the corresponding nested sequent
systems~\cite{str:fossacs13,marin:str:aiml}. It is possible to express
Maehara style multi-conclusion systems in nested
sequents~\cite{str:2017maehara}, and therefore also in ordinary
labelled sequents. However, also in these systems there are rules
($\rlabrn\IMP$ and $\rlabrn\BOX$) that force a single-conclusion
premise, even though this is not the case in labelled
systems~\cite{negri:jpl2005} or nested sequents~\cite{fitting:83}\lutz{wrong ref} for
intuitionistic logic.

### 2020-09-30 FullyLabelled/JLC/main.tex
bugfixes in 1 and 2

 \begin{abstract}
   %% \review{Reconsider abstract. It reads very “operational” and does not clearly state the contributions of the paper.}
   %% \lutz{Done?}
   %% \sonia{Added two sentences at the beginning and modified it a bit.}
   Labelled proof theory has been famously successful for modal logics by mimicking their relational semantics within deductive systems. Simpson in particular designed a framework to study a variety of intuitionistic modal logics integrating a binary relation symbol in the syntax.
   %%
   In this paper, we present a labelled sequent system for
   intuitionistic modal logics such that there is not only one, but
   two relation symbols appearing in sequents: one for the accessibility
   relation associated with the Kripke semantics for normal modal logics and
   one for the preorder relation associated with the Kripke semantics
   for intuitionistic logic.
   %%
   This puts our system in close correspondence with the standard
   birelational Kripke semantics for intuitionistic modal logics. 
   %%
   As a consequence it can be extended with arbitrary intuitionistic Scott-Lemmon axioms. 
   %%
   We show soundness and completeness, together with an internal cut elimination proof, encompassing a wider array of intuitionistic modal logics than any existing labelled systems.
 \end{abstract}

>>>>

 \begin{abstract}
   %% \review{Reconsider abstract. It reads very “operational” and does not clearly state the contributions of the paper.}
   %% \lutz{Done?}
   %% \sonia{Added two sentences at the beginning and modified it a bit.}
   Labelled proof theory has been famously successful for modal logics by mimicking their relational semantics within deductive systems. Simpson in particular designed a framework to study a variety of intuitionistic modal logics integrating a binary relation symbol in the syntax.
   %%
   In this paper, we present a labelled sequent system for
   intuitionistic modal logics such that there is not only one, but
   two relation symbols appearing in sequents: one for the accessibility
   relation associated with the Kripke semantics for normal modal logics and
   one for the preorder relation associated with the Kripke semantics
   for intuitionistic logic.
   %%
   This puts our system in close correspondence with the standard
   birelational Kripke semantics for intuitionistic modal logics. 
   %%
   As a consequence it can be extended with arbitrary intuitionistic Scott-Lemmon axioms. 
   %%
   We show soundness and completeness, together with an internal cut elimination proof, encompassing a wider array of intuitionistic modal logics than any existing labelled system.
 \end{abstract}

### 2020-09-30 FullyLabelled/JLC/preliminaries.tex
bugfixes in 1 and 2

The language of {intuitionisitic modal logic} is the one of intuitionistic propositional logic with the modal operators $\BOX$ and $\DIA$, standing most generally for \emph{necessity} and \emph{possibility}.
%
Starting with a set $\mathcal{A}$ of atomic propositions, denoted by lower case letters $\fm a$, $\fm b$, $\fm c$,\ldots, modal formulas, denoted by capital letters $\fm A$, $\fm B$, $\fm C$, \ldots, are constructed from the grammar:
%
$$
\fm A \coloncolonequals
\fm a \mid \fm{A \AND A} \mid \fm{A \OR A} \mid \fm \BOT \mid \fm{A \IMP A} \mid \fm{\BOX A} \mid \fm{\DIA A}
$$
%

@@

		\item[($\rn{F_2}$)] For all $\lb {u'}, \lb u, \lb v \in W$, if $\futs uv$, there exists $\lb {v'}$ s.t.~$\accs{u'}{v'}$ and $\futs v{v'}$.

@@

\begin{definition}
	\label{model}
	A \emph{bi-relational model} $\M$ is a quadruple $\langle W, \rel,\le,V \rangle$ with $\langle W, \rel, \le \rangle$ a bi-relational frame and $V\colon W \to 2^\mathcal{A}$ a monotone valuation function, that is, a function mapping each world $\lb w$ to the subset of propositional atoms true at $\lb w$, additionally subject to:
	if $\lb w \le \lb{w'}$ then $V(\lb w)\subseteq V(\lb{w'})$.
\end{definition}

@@

		\M, \lb w \Vdash \fm{\BOX A} & \mbox{iff} & \mbox{for all } \lb{w'} \mbox{ and } \lb u \mbox{ with } \futs w{w'} \mbox{ and } \accs {w'}u, \M, \lb u \Vdash \fm A \hfill \\ %$(\ast)$

@@

From the monotonicity of the valuation function $V$, we get a monotonicity property for the relation:
\begin{proposition}(Monotonicity) 
	For any formula $\fm A$ and for $\lb w, \lb{w'} \in W$, if $\futs w{w'}$ and $\M, \lb w \Vdash \fm A$, then $\M, \lb{w'} \Vdash \fm A$.
\end{proposition}

>>>>

The language of {intuitionisitic modal logic} is the one of intuitionistic propositional logic with the modal operators $\BOX$ and $\DIA$, standing most generally for \emph{necessity} and \emph{possibility}.
%
Starting with a set $\mathcal{A}$ of atomic propositions, denoted by lower case letters $\fm a$, $\fm b$, $\fm c$, \ldots, modal formulas, denoted by capital letters $\fm A$, $\fm B$, $\fm C$, \ldots, are constructed from the grammar:
%
$$
\fm A \coloncolonequals
\fm a \mid \fm{A \AND A} \mid \fm{A \OR A} \mid \fm \BOT \mid \fm{A \IMP A} \mid \fm{\BOX A} \mid \fm{\DIA A}
$$
%

@@

		\item[($\rn{F_2}$)] For all $\lb {u'}, \lb u, \lb v \in W$, if $\futs u{u'}$ and $\accs uv$, there exists $\lb {v'}$ s.t.~$\accs{u'}{v'}$ and $\futs v{v'}$.

@@

\begin{definition}
	\label{model}
	A \emph{bi-relational model} $\M$ is a quadruple $\langle W, \rel,\le,V \rangle$ with $\langle W, \rel, \le \rangle$ a bi-relational frame and $V\colon W \to 2^\mathcal{A}$ a monotone valuation function, that is, a function mapping each world $\lb w$ to the subset of propositional atoms that are true at $\lb w$, additionally subject to:
	if $\lb w \le \lb{w'}$ then $V(\lb w)\subseteq V(\lb{w'})$.
\end{definition}

@@

		\M, \lb w \Vdash \fm{\BOX A} & \mbox{iff} & \mbox{for all } \lb{w'} \mbox{ and } \lb u \mbox{ with } \futs w{w'} \mbox{ and } \accs {w'}u \mbox{, we have } \M, \lb u \Vdash \fm A \hfill \\ %$(\ast)$

@@

From the monotonicity of the valuation function $V$, we get a monotonicity property for the $\Vdash$ relation:
\begin{proposition}(Monotonicity) 
	For any formula $\fm A$ and for any $\lb w, \lb{w'} \in W$, if $\futs w{w'}$ and $\M, \lb w \Vdash \fm A$, then $\M, \lb{w'} \Vdash \fm A$.
\end{proposition}
